---
title: "Homework_7_Jared_Lee"
author: "Jared Lee"
date: "7/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(broom)
library(readr)
library('rvest')
library(xml2)
library(dbplyr)
library(knitr)
library(tidyr)
library(readr)
```

##Exercise 1
###Load the train.csv dataset into R. How many observations and columns are there?
1460 observations. 81 variables
```{r} 
file_path <- '/Users/jalee/downloads/train.csv'
train_data <- read_csv(file = file_path)

glimpse(train_data)
```


##Exercise 2
###Our target will be SalePrice.

####Visualize the distribution of SalePrice.
```{r}
train_data %>% 
  ggplot(aes(SalePrice)) +
  geom_histogram()
```

####Visualize the covariation between SalePrice and Neighborhood.
```{r}
train_data %>%
  ggplot() +
  geom_point(aes(SalePrice, Neighborhood))
```


####Visualize the covariation between SalePrice and OverallQual.
```{r}

```

##Exercise 3
###take a look at the coefficient,
###compare the coefficient to the average value of SalePrice, and
###take a look at the R-squared.

##Exercise 4


##Exercise 5 (OPTIONAL)


##Exercise 6