---
title: "Homework_6_Jared_Lee"
author: "Jared Lee"
date: "7/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(mosaicData)
library(ggplot2)
```
##Exercise 1
###1) What variables are in this data set?
Answer: outcome, smoker, age

###2) How many observations are there and what does each represent?
Answer: 1,314 obervations representing individuals regarding survival status (dead, alive), smoker or nonsmoker, and age in years.

###3) Create a table (use the R code below as a guide) and a visualization of the relationship between smoking status and outcome, ignoring age. What do you see? Does it make sense?
Answer: I would expect the number of deaths to be higher for smokers. Non-smoker death proportion is 31% vs 24% for smokers. I would say that it does not make sense.

```{r}
Whickham %>% 
  count(smoker, outcome)

Whickham %>% 
  count(smoker, outcome) %>%
  ggplot() +
  geom_point(aes(smoker, n, color = outcome)) +
  ylab("count")

```

###4)Recode the age variable into an ordered factor with three categories: age <= 44, age > 44 & age <= 64, and age > 64. Now, recreate visualization from above, but facet on your new age factor. What do you see? Does it make sense?
Answer: The number of deaths for the 65+ age group exceeds the alive numbers. This makes more sense as health symptoms may be accelorate a decline in health later in life.

```{r}
Whickham_age <- Whickham %>% mutate(grp_age = case_when(age <= 44 ~ '44 or less',
                                                      age > 44 & age <= 64 ~ '45 to 64',
                                                      age > 64 ~ '65+'))

Whickham_age %>% 
  count(smoker, outcome, grp_age) %>%
  ggplot() +
  geom_point(aes(smoker, n, color = outcome)) +
  facet_wrap(~grp_age)

```

##Exercise 2
###1) Generate a random sample of size n = 10000 from a gamma(1,2) distribution and plot a histogram or density curve. Use the code below to help you get your sample.
```{r}
n <- 10000

# look at ?rgamma to read about this function
(gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2)))

gamma_samp %>% 
  ggplot() + 
  geom_histogram(aes(x))

```


###2) What is the mean and standard deviation of your sample? They should both be close to 2 because for a gamma distribution:
```{r}
(mean_samp <- gamma_samp %>% .[['x']] %>% mean())

(sd_samp <- gamma_samp %>% .[['x']] %>% sd())
```


###3) Pretend the distribution of our population of data looks like the plot above. Now take a sample of size n = 30 from a Gamma(1,2) distribution, plot the histogram or density curve, and calculate the mean and standard deviation.

```{r}
samp30 <- gamma_samp %>% 
  sample_n(30)

samp30 %>%
  ggplot() + 
  geom_histogram(aes(x))

(mean_samp1 <- samp30 %>% .[['x']] %>% mean())

(sd_samp1 <- samp30 %>% .[['x']] %>% sd())  
  
```


###4) Take a sample of size n = 30, again from the Gamma(1,2) distribution, calculate the mean, and assign it to a vector named mean_samp. Repeat this 10000 times!!!! The code below might help.

```{r}
# create a vector with 10000 NAs
mean_samp2 <- rep(NA, 10000)

# start a loop
for(i in 1:10000) {
  g_samp <- rgamma(30, shape = 1, scale = 2)
  mean_samp2[i] <- mean(g_samp)
}
# Convert vector to a tibble
(mean_samp2 <- tibble(mean_samp2))

View(mean_samp2)
```


###5) Make a histogram of your collection of means from above (mean_samp).
```{r}
mean_samp2 %>%
  ggplot() +
  geom_histogram((aes(mean_samp2)))


```

###6) Calculate the mean and standard deviation of all of your sample means.

```{r}
(mean_a <- mean_samp2 %>% .[['mean_samp2']] %>% mean())

(sd_a <- mean_samp2 %>% .[['mean_samp2']] %>% sd())  
```


###7) Did anything surprise you about your answers to #6?
Answer: The SD is very low. Eariler it was closer to 2

###8) Repeat #4-#6, but now with a sample of size n = 300 instead. Do your results match up well with the theorem?
Answer: Distribution is still normal.
```{r}
# create a vector with 10000 NAs
mean_samp2 <- rep(NA, 10000)

# start a loop
for(i in 1:10000) {
  g_samp <- rgamma(300, shape = 1, scale = 2)
  mean_samp2[i] <- mean(g_samp)
}

# Convert vector to a tibble
(mean_samp2 <- tibble(mean_samp2))

View(mean_samp2)

#Histogram
mean_samp2 %>%
  ggplot() +
  geom_histogram((aes(mean_samp2)))

#Find Mean and SD
(mean_a <- mean_samp2 %>% .[['mean_samp2']] %>% mean())

(sd_a <- mean_samp2 %>% .[['mean_samp2']] %>% sd())  
```



