---
title: "COMPSCIX 415.2 Homework 1"
author: "Jared Lee"
date: "6/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: https://github.com/jar1818/compscix-415-2-assignments.git 

```{r warning=FALSE, message=FALSE}
library(mdsr)
```

```{r}
library(tidyverse)
```
  
```{r load_packages, warning=FALSE, message=FALSE}
data("WorldCities")
```

# QUESTION 1

View sample of the data WorldCities
```{r}
glimpse(WorldCities)
```

There are **23,018** observations in the WorldCities data. Some of the variable names are *code*, *name*, *latitude*, *longitude*, and *country*.

View 200 obersrvations for all varibles 
```{r}
WorldCities <- head(WorldCities, 200)
```

View country obervations
```{r}
country_col <- WorldCities$country
```

View distinct countries
```{r}
unique(country_col)
```

# QUESTION 2

### Compile and commit
I made some style changes (bold, italics) and removed warnings above. 

View country observations using Select
```{r}
country_col <- select(WorldCities, country)
```

View country observations using %>% and Select
```{r}
country_col <- WorldCities %>% select(country)
```

View top 5 overservations of Region variable
```{r}
WorldCities %>% select(region) %>% head(5)
```


# QUESTION 3

View distinct Countries using two pipe operators
```{r}
WorldCities %>% select(country) %>% unique()
```

Plot population observations
```{r}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram()
```

Change X axis label to 'City Population'
```{r}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram() + xlab('City Population')
```

# Question 4

Plot population with new x label, new y label, new title, and white background
```{r}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram() + ylab('y axis') + xlab('x axis') + ggtitle('my title') + theme_bw()
```

