---
title: "homework_2_Lee_Jared"
author: "Jared Lee"
date: "6/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(mdsr)
```


## Section 3.2.4: all exercises
### 1) Run ggplot(data = mpg). What do you see?

Answer: ggplot(data = mpg) shows an empty graph.
```{r}
ggplot(data = mpg)
```

### 2) How many rows are in mpg? How many columns?
Answer: 
rows: 234
columns: 11
```{r}
glimpse(mpg)
```

### 3) What does the drv variable describe? Read the help for ?mpg to find out.
Answer: 
drv describes the drive train.
f = front-wheel drive, r = rear wheel drive, 4 = 4wd
```{r}
?mpg
```

### 4) Make a scatterplot of hwy vs cyl.
Answer: 
```{r}
ggplot(mpg) +
  geom_point(aes(x = hwy, y = cyl))
```

### 5) What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
Answer: 
Both variables are descrete, so showing a table or raw numbers is a better visualization. 
```{r}
ggplot(mpg) +
  geom_point(aes(class, drv))
```


##Section 3.3.1: all exercises
### 1) What’s gone wrong with this code? Why are the points not blue?

Since color = "blue" is inside the aesthetic function, R treats it like a dimension to be added as opposed to a color treatment. This should be outside of aes() instead.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

### 2) Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

manufacturer: categorical,

displ: continuous,

year: categorical,

cyl: categorical,

trans: categorical,

drv: categorical,

cty: continuous,

hwy: continuous,

fl: categorical,

class: categorical

```{r}
glimpse(mpg)
```


### 3) Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **

For color, continuous variables show a gradient. For categorical, each variable has their own unique color. 
Similarly with size, continuous variables are a scale of size from small to large. And categorical variables get their own unique size. 

For Shape, continuous variables cannot be used as there is no scale for shapes.
```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy, shape = class))
```

### 4) What happens if you map the same variable to multiple aesthetics?
It creates a linear line. 

```{r}
ggplot(mpg) + 
  geom_point(aes(x = hwy, y = hwy))
```


### 5) What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

The stroke aesthetic defines the width or thickness of the border of the points. 
```{r}
?geom_point
```


```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy, shape = class, stroke = 5))
```



### 6) What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?
A boolean is created. 
```{r}
ggplot(mpg) +
  geom_point(aes(x = hwy, y = class, colour = drv < 5))
```


##Section 3.5.1: #4 and #5 only

### 4) Take the first faceted plot in this section. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? 
####Advantage: 
It's visually easier to see observations by isolating them on a separate graph(s). 
####Disadvantage: 
There can be too many graphs where the viewer doesn't know where to look. Also, it's difficult to compare behaviors of one descrete variable to another with so many graphs. 
#### How might the balance change if you had a larger dataset?
With a large dataset, there may be too many graphs to look at and comparative analysis would be difficult.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

###5) Read ?facet_wrap. What does nrow do? What does ncol do? 

nrow is how you define the number of rows
ncol is how you define the number of columns

####What other options control the layout of the individual panels? 
scales - set the scales of the axises to fixed or free
shrink - shrinks the scales to fit the statistics
labeller - A function that takes one data frame of labels and returns a list or data frame of character vectors
as.table - facets are laid out like a table with highest values at the bottom-right.
switch - choose which graph contains the lables

####Why doesn’t facet_grid() have nrow and ncol argument?
With facet_grid(), rows and columns should be defined by the number of unique values for that variable.

```{r}
?facet_wrap
```


##Section 3.6.1: #1-5. Extra Credit: Do #6
###1) What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
Line chart: geom_line()
Boxplot: geom_boxplot()
Histogram: geom_histogram()
Area Chart: geom_area()

###2)Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
Prediction: Two different geommetrical funtions are used. The x and y axises are defined in ggplot() and apply to both. Color that adds a 3rd variable is also added here so both geom funtions should be color coded by drv.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
Prediction check: se = FALSE controls whether the grey bands are displayed which show the variation. 

###3)What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?
show.legend = FALSE removes the legend from the graph. If you remove that line, the legend appears which is the default. I'm guessing it was used earlier in the chapter to create a simpler illustration of how geom_smooth() works.
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, color = drv), show.legend = FALSE)
```

###4 What does the se argument to geom_smooth() do?
The se argument displays grey area to show the distribution of the observations.

###5) Will these two graphs look different? Why/why not?
The two graphs look the same. The x and y axises are just defined differently. The first one sets x and y globally. The second sets x and y for each geometric funtion, but they are mapped the same. 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

###6) 
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(stroke = 2) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(group = drv), se = FALSE) +
  geom_point(stroke = 2)

```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(se = FALSE) +
  geom_point(stroke = 2)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(se = FALSE, size = 2) +
  geom_point(aes(color = drv), stroke = 2)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(linetype = drv), se = FALSE, size = 2) +
  geom_point(aes(color = drv), stroke = 2)
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, fill = drv)) +
  geom_point(stroke = 2, shape = 21, color = 'white')
```


##Section 3.7.1: #2 only
###2) What does geom_col() do? How is it different to geom_bar()?
Geom_col creates a stacked bar where you have to define y. Each grouping of y is represented by a different color/shade. Geom_bar uses a count rather than defining y. 

```{r}
ggplot(diamonds) +
  geom_col(aes(x = cut, y = price))
```

```{r}
ggplot(diamonds) +
  geom_bar(aes(x = cut))
```


## Extra Credit: Look at the data graphics at the following link: What is a Data Scientist. Please briefly critique the designer’s choices. What works? What doesn’t work? What would you have done differently?
 
####What works:
- The heading for each section states the key finding from the visualizations below them. 
- The graphs/charts are reasonably simplistic and don't show overwhelming amounts of data. 

####What doesn't work:
- Donut chart is difficult to compare pieces/slices to each other. Small sections aren't represented clearly. 
- The area blocks are difficult to distingush. The values are so close to eachother, it doesn't make sense to visualize them this way when they're so similar.
- The bar charts show far too many charts. Grouping the data into 1 or 2 charts/graphs would be easier to read and compare than 10 invidual charts. 

####Do differently:
- All visualizations are a shade of blue. Another color to highlight key insights would be extremely helpful.
- The top of the infographic doesn't define the problem well. It's not summarized at the top what the study is about. Having just a paragraph is too verbose.
- The end could use a summary of the key learnings. 





