---
title: "homework_3_Lee_Jared"
author: "Jared Lee"
date: "6/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Loading Packages
```{r}
library(tidyverse)
library(nycflights13)
library(dplyr)
```


##Section 3.8.1: all exercises
### 1) What is the problem with this plot? How could you improve it?
Answer: There are overlapping points which hides the number of actual values. By using the geom_jitter function, you can see all of the points plotted.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()
```

###2) What parameters to geom_jitter() control the amount of jittering?
Answer: Width and Height control the amount of jittering

```{r}
?geom_jitter()
```

###3) Compare and contrast geom_jitter() with geom_count().
Answer: geom_jitter() will plot each individual point separately. geom_count() will show an increased size for points that have higher frequency. 
```{r}
?geom_count()
```

###4) What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.
Answer: REVISIT THIS!! (dodge is default)
```{r}
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x = drv, y = hwy), position = "identity")
```


##Section 3.9.1: #2 and #4 only
### 2) What does labs() do? Read the documentation.
Answer: Allows customization of axis and legend labels. 

### 4) What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?
Answer: Highway miles always exceeds City miles but not a perfect ratio does not apply to all cars. 
coord_fixed() specifies the relationship between the x and y coordinates so the marks on the axises have a consistant spacing and number values. 
geom_abline() displays a line specified by slope and intercept.

##Section 4.4: #1 and #2 only 

###1) Why does this code not work?
Answer: The "ı" character should be an "i"

###2) Tweak each of the following R commands so that they run correctly:
Answer:
```{r}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```



##Section 5.2.4: #1, #3 and #4 only. You will need to install the nycflights13 package and use the flights data.

##1) Find all flights that

###a) Had an arrival delay of two or more hours
```{r}
(a <- filter(flights, arr_delay >= 120))
```


###b) Flew to Houston (IAH or HOU)
```{r}
(b <- filter(flights, dest %in% c("IAH", "HOU")))
```


###c) Were operated by United, American, or Delta
```{r}
(c <- filter(flights, carrier %in% c("UA", "AA", "DL")))
```

###d) Departed in summer (July, August, and September)
```{r}
(d <- filter(flights, month %in% c(7,8,9)))
```


###e) Arrived more than two hours late, but didn’t leave late
```{r}
(e <- filter(flights, arr_delay > 120, dep_delay <= 0))
```


###f) Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
(f <- filter(flights, dep_delay >= 60, arr_delay <= dep_delay - 30))
```


###g) Departed between midnight and 6am (inclusive)
```{r}
(g <- filter(flights, dep_time >= 000 & dep_time <= 600))
```


##3) How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
Answer: 8,255 flights missing a dep_time. dep_delay, arr_time, arr_delay, tailnum, and air_time have NA values also. They may represent cancelled flights. 
```{r}
filter(flights, is.na(dep_time))
```


##4) Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
Answer: Any number to the zero power is 1. 

##Section 5.4.1: #1 and #3 only

###1)Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```
```{r}
select(flights, starts_with("dep"), starts_with("arr"))
```
```{r}
select(flights, dep_time:arr_delay, -sched_dep_time, -sched_arr_time)
```


###3)What does the one_of() function do? Why might it be helpful in conjunction with this vector?



